#Loading and preprocessing the data

1. Load the data (i.e. read.csv())
2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
if(!file.exists("getdata-projectfiles-UCI HAR Dataset.zip")) {
        temp <- tempfile()
        download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
        unzip(temp)
        unlink(temp)
}

data <- read.csv("activity.csv")
```

#What is the mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
```{r}
steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
rmean <- mean(steps_by_day$steps)
rmedian <- median(steps_by_day$steps)
```
The mean of the total number of steps taken per day is `r rmean` and the median is `r rmedian`.

#What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
steps_by_interval <- aggregate(steps ~ interval, data, mean)

plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
```
The 5-minute interval, on average across all the days in the data set, containing the maximum number of steps is `r max_interval`.

#Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
incomplete<-sum(is.na(data$steps))
```
The total number of missing values in the dataset is `r incomplete`.

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Missing values were imputed by inserting the average for each interval. 

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), steps_by_interval$steps[match(data$interval, steps_by_interval$interval)], data$steps))
```

4. Make a histogram of the total number of steps taken each day.
```{r}
steps_by_day_i <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_by_day_i$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "red"), lwd=10)
```

Calculate and report the mean and median total number of steps taken per day. 
```{r}
rmean.i <- mean(steps_by_day_i$steps)
rmedian.i <- median(steps_by_day_i$steps)
```
The mean total number of steps taken per day based on inputed data is `r rmean.i`.
The median total number of steps taken per day based on inputed data is `r rmedian.i`.

Do these values differ from the estimates from the first part of the assignment? 
```{r}
mean_diff <- rmean.i - rmean
med_diff <- rmedian.i - rmedian
```
The difference between the non-imputed and imputed means is `r mean_diff`.
The difference between the non-imputed and imputed medians is `r med_diff`.

What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
total_diff <- sum(steps_by_day_i$steps) - sum(steps_by_day$steps)
```
There were `r total_diff` more steps in the inputed data.

#Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}
steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```